Integrating Social Visual Analysis and Deployment with RCloud

### 2015-03-30 discussion.tex
pare down interviews more, discuss limitations raised

\gordon{This section now overlaps with the Interview Study quite a bit - we can
  use that to motivate this, but that would mean not immediately responding to
  the complaints raised there. On the other hand, trying to merge these two
  sections would be a beast. The good news is, we anticipated most of the
  problems that people are facing now. :-/}

@@

{\bf Shared artifacts and artifact histories} are a central feature
of RCloud, through its shared workbooks of experiments and analyses.
We observed that hackers readily shared work through RCloud. They
used it to discuss ideas and results in data analysis, to ask for help
with programming problems, and to publish results by sharing URLs.
Artifact histories are provided through a version management system (git).
\stephen{Do analysts type raw git commands?}

@@

{\bf Multiple language support.}
We did not foresee the demand to relax the assumption that programming
is done mainly in R. In retrospect, it is clear that many data science
problems are solved with cross-language approaches (for example, scikit-learn
in Python, weka in Java).\stephen{Isn't this fixed in some way now?}
\gordon{Should we have a section on ways that we've responded to popular demand?}

{\bf Versioning.}
Some important aspects of the environment are more difficult to manage
in RCloud than in conventional systems. RCloud does not explicitly
separate the development and deployment environments. More than that,
ever version of every prototype is shared by default. Although this encourages
collaboration on work in progress it also quickly exposes errors such as
misconfiguration, mismatches between versions of packages, and programming
errors that can unintentionally affect production websites. Such errors
are difficult to completely avoid, but the power of convenient sharing is often
worth it (as proven true with distributed systems in general).
Still, effective control over versioning when sharing problem needs more
attention if RCloud would be scaled up to a wider number of users lacking
informal channels to coordinate work.
Versioning needs to be managed in several places: in scripts;
in the R environment such as the installed libraries and packages;
and in the external environment such as the operating system and
protocols spoken by remote services. At one end of the scale, we have
full control over the versioning of scripts via git, along with conventions
to name stable or working versions of scripts, and versions known to work together.
The R environment itself is under the control of its package manager
that has rules to ensure reasonable consistency.
It is at least possible for RCloud to access information about
package versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not strong.
At the other end of the scale, there is not much reason to expect version
control for the external environment. Most programmers rely on clean living
and a careful approach to system upgrades.

>>>>

@@

{\bf Shared artifacts and artifact histories} are a central feature of
RCloud, through its shared workbooks of experiments and analyses.  We
observed that hackers readily shared work through RCloud. They used it to
discuss ideas and results in data analysis, to ask for help with
programming problems, and to publish results by sharing URLs.  Artifact
histories can be accessed through the notebook tree or through GitHub's web
interface.

@@

{\bf Exploration is hard.}
R is a popular language for exploratory data analysis, and we built RCloud
as a way for data scientists to seamlessly transition from EDA to sharing
and presentation. But the interface we implemented for combining sharing and presentation
into the R command line interferes with the exploration aspect.

In existing R tools, the command prompt is used to try things out. It is very
fast and there is no commitment: nothing gets saved and there is nothing to
later clean up. In contrast, in RCloud's notebook interface, the extra cells need
to get deleted later, and there seems to be a mental burden that ``bad stuff
is getting saved.''

One thing to try here is making deletion easier, for example with a
shortcut key to remove the last cell, or a way to keep a few cells and cull
the rest. A far more ambitious solution, suggested by one of our
interviewees, is to implement auditing of data analyses as it existed in
S \cite{Becker:1988:Auditing}, so that the code which did not lead to a
result could be discarded automatically.

The interface is also perceived to be slower, both in terms of reaction
time---RCloud necessarily saves a cell to its repository before executing
it---and because it requires mouse interaction.

Other complaints mostly focus on layout and can probably be addressed with
some iteration and customization features.

{\bf Discovery is hard.}
While we have provided tools for searching and curating notebooks, the
search function does not help for finding methodology (and does not protect
searchers from erroneous notebooks), and the notebook tree starts to get
unwieldly with hundreds of users with hundreds of notebooks. In addition,
our choice to retain notebooks even when they have been deleted has
exacerbated the situation, since obsolete and broken notebooks stay around
forever.

One thing to try is tagging and filtering the tree on tags, so that users
can mark their own notebooks or those of others to make them easier to
find.

However, this relies on users making a conscious effort, and it may be
better to rely more on passive metrics such as how often a notebook has
been run.

The system also needs to help users curate their notebooks. If a notebook
has been superceded by one of its forks, a user who opens it should be
warned. If a notebook has been deleted, but someone still has a link to it,
it should be clear that the notebook is dead. And operations on multiple
notebooks and folders of notebooks should be supported to make curation
easier.

{\bf Collaboration is hard.}
While forking provides a simple way to continue someone else's work, working
in teams is a lot more complicated. The forking feature is popular but it
poses a problem, too, as dozens of versions of a notebook start proliferating.

One solution we are mulling over is to allow overwriting the current state
of a notebook with the current state of another fork of the notebook. More
general solutions would be serious engineering endeavors: even GitHub's web
interface does not support merging where there are any conflicts. Actual
sharing of notebooks would probably require moving toward git's notion of
decentralized code repositories, because GitHub gists do not support shared
ownership.

{\bf Versioning.}
Some important aspects of the environment are more difficult to manage
in RCloud than in conventional systems. RCloud does not explicitly
separate the development and deployment environments. More than that,
ever version of every prototype is shared by default. Although this encourages
collaboration on work in progress it also quickly exposes errors such as
misconfiguration, mismatches between versions of packages, and programming
errors that can unintentionally affect production websites. Such errors
are difficult to completely avoid, but the power of convenient sharing is often
worth it (as proven true with distributed systems in general).
Still, effective control over versioning when sharing problem needs more
attention if RCloud would be scaled up to a wider number of users lacking
informal channels to coordinate work.

Versioning needs to be managed in several places: in scripts;
in the R environment such as the installed libraries and packages;
and in the external environment such as the operating system and
protocols spoken by remote services. At one end of the scale, we have
full control over the versioning of scripts via git, along with conventions
to name stable or working versions of scripts, and versions known to work together.
The R environment itself is under the control of its package manager
that has rules to ensure reasonable consistency.
It is at least possible for RCloud to access information about
package versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not strong.
At the other end of the scale, there is not much reason to expect version
control for the external environment. Most programmers rely on clean living
and a careful approach to system upgrades.

### 2015-03-30 interviews.tex
pare down interviews more, discuss limitations raised

Taylor develops packages for analysts: ``I use it more for sharing code with
other people, and for doing tutorials for iotools or hmr'', the packages he
maintains. ``I want people to see how the package works, so I clearly want them
to see the code [...] I kind of write it just like I would write GitHub
Markdown, where you have little code snippets and text, but RCloud lets me
actually run the snippets'' and display the results. He also uses RCloud for
explaining issues with data sources.

@@

Forking also provides a way for others to troubleshoot when something goes
wrong.  When Horace works with the users of his notebooks, ``I'll teach people
to intercept the result in the middle,'' to ``insert print statements here and
there and check values''

@@

Currently parameters can be added to the URL, but adding user interface elements
to do the same thing requires expertise. The tool-builders see a need for better
ways to parameterize notebooks. Taylor complains that a notebook might say ``
`This is a report of the volume of all of our feeds for this month', and someone
would want to look at it for the next month or the previous month, so they'd
fork it to change the month.''

Jyothi calls it ``web-enabling'' the notebook: users ``can always fork the
notebooks and make changes, but I feel that if the owner of the notebook
web-enables it [...] it's easier to run. Instead of
forking it, if they can set options, it's probably more efficient, and also they
can [still] fork it if they want to.''

@@

\subsection{Discovering others' work}
Many users find the search function useful. Raif says ``the fact that we have
all the notebooks there, searchable, saves me from replicating what other people
have done. [...] One way of thinking of it is like Google Code: you just search
for something and you find something useful and you just use it.''

@@

\subsection{Web integration}
Integrating an analysis language into a web development environment is something
the tool developers really appreciate.  The structure of Horace's visualization
notebook means
\begin{quote}
the other guys who want to do analytics on the data, they can first pull the
data, do the analytics on it, and then feed the viewer the data. So [...]
anyone from the stats group [...] can always insert something in between. We
were trying to make such an example so that they don't have to rewrite the GUI
part and they know how to get the data source. Once you get the data into
RCloud, then you have a dataframe to work with, and then they know how to
produce another dataframe.
\end{quote}

@@

The cell structure also can be problematic if some cells take a long time to
run. In Kenny's work, there is often a ``long tail''.  The first cell may take
seconds to run, the next cell minutes, and the last cell 5 hours.  In this
situation, the ``Run whole notebook'' button is ``dangerous''. When converting
his work to notebooks, Kenny ends up with a lot of comments that say ``this cell
takes a long time to run.'' This is also a huge
issue for demoing code.

The solution is probably something like knitr's caching of results, which Kenny
is doing in an ad-hoc way, ``littering the notebook with little switches that
comment out'' the slow parts and ``load the object instead'', ``save the object
to disk''. One of the authors of this paper manually writes cells that check if
a result file exists, and perform the calculation only if it's not there.

Rick on auditing (possible solution): ``you didn't see your mistakes, they were
in there but you never really saw them because a mistake never led to an answer
at the end'' ``the real result of this analysis was these three plots, so go
back and figure out everything that I did that was involved in creating those
three plots. So that I could start from the raw data and create those three
plots.''

@@

Taylor, who writes tools and example notebooks for those tools, seems to be
haunted the most by dead notebooks.  At first, he had to learn how to write safe
notebooks: ``It would be a file inside a directory that only I had [...] access
to and so someone else couldn't run it, [...] a temporary file on something that
was read-only to everyone else. But I've gotten careful about that.''

But even correcting notebooks can come too late:
\begin{quote}
I'll make a notebook of a
little tutorial of basic functions and what they do, and then I'll send [it]
out, and then people will give me feedback on [it], or have some new feature, or
I'll realize something's wrong, and then I'll go back and change it and update
that page, but then what happens is people have in the meantime forked it. [...]
Usually what it'll be is actually that I'll have to change a package as well, so
their old fork stops working, and then they complain. [...] This probably bites
me more than other people, because I'm doing development so things are changing
as they're looking at the code. Like, iotools, a lot of the functions change
names or change syntax.
\end{quote}

Taylor tried to keep his notebook tree tidy, but this didn't help either, because
notebooks still exist when they have left the tree.
\begin{quote}
That's just a problem because people will have old notebooks that
don't work anymore. It's annoying when it's a notebook that I still have and the
format has changed just a bit. More than that is when people have, now that I
know that you can't even delete notebooks, they'll have a fork of something that
doesn't exist anymore on my thing [...] Now I'm really gun-shy about sharing
notebooks because it's like, do I want to support this forever?

[...] A good example is, someone was complaining about doing forking in an
RCloud notebook. They were trying to use a parallel package. And I said, no no I
think it works, and I did like a one-cell thing that just did some sweep kind of
thing to show that it was actually working in parallel, and then I just deleted
it, thinking it was gone. And then it saved something in a directory that I then
deleted, because it was just a temporary directory. Three months later, I don't
know why they were running it, they said `oh it doesn't work'. I think that's
the biggest issue I'm having right now, all these forks.
\end{quote}

Taylor and Horace work together on a notebook.

\begin{quote}
There's no way for both of us
to have ownership of a notebook, so the only way is to fork it back and forth,
and so we have dozens of old copies. [We end up] deleting all the old ones, but
people still have links to them, because they don't actually disappear. [...]
The UI makes it easy to go back to the old stale versions, because you can say
`forked from' `forked from' `forked from', but there's actually no way to go the
way that we want people to go, which is, like, up the tree.
\end{quote}

>>>>

Taylor develops packages for analysts, and uses RCloud ``more for sharing code with
other people, and for doing tutorials for iotools or hmr'', his packages.
``I want people to see how the package works, so I clearly want them
to see the code [...] I kind of write it just like I would write GitHub
Markdown, where you have little code snippets and text, but RCloud lets me
actually run the snippets'' and display the results. He also uses RCloud for
explaining problems with data sources.

@@

Forking also provides a way for others to troubleshoot when something goes
wrong.  When Horace works with the users of his notebooks, ``I'll teach people
to intercept the result in the middle,'' to ``insert print statements here and
there and check values.''

@@

Currently parameters can be added to the URL, but adding user interface elements
to do the same thing requires expertise. The tool-builders see a need for
facilities to make it easier. Taylor complains that a notebook might say ``
`This is a report of the volume of all of our feeds for this month', and someone
would want to look at it for the next month or the previous month, so they'd
fork it to change the month.''

Jyothi calls it ``web-enabling'' the notebook: users ``can always fork the
notebooks and make changes, but I feel that if the owner of the notebook
web-enables it, [...] it's easier to run. Instead of
forking it, if they can set options, it's probably more efficient, and also they
can [still] fork it if they want to.''

@@

\subsection{Discovering others' work}
Many users find the search function useful. Raif says ``the fact that we have
all the notebooks there, searchable, saves me from replicating what other people
have done.''

@@

\subsection{Integrated analysis}
Integrating an analysis language into a web development environment is something
the tool developers really appreciate.  The structure of Horace's visualization
notebook means
\begin{quote}
the other guys who want to do analytics on the data, they can first pull the
data, do the analytics on it, and then feed the viewer the data. So [...]
anyone from the stats group [...] can always insert something in between. We
were trying to make such an example so that they don't have to rewrite the GUI
part and they know how to get the data source. Once you get the data into
RCloud, then you have a dataframe to work with, and then they know how to
produce another dataframe.
\end{quote}

@@

The cell structure also can be problematic if some cells take a long time to
run. In Kenny's work, there is often a ``long tail''.  The first cell may take
seconds to run, the next cell minutes, and the last cell 5 hours.  In this
situation, the run whole notebook button is ``dangerous''. When converting
his work to notebooks, Kenny ends up with a lot of comments that say ``this cell
takes a long time to run.''

To avert this danger, users are forced to write code within the notebook to
cache results. Kenny ends up ``littering the notebook with little switches that
comment out the slow parts'' and either ``load the object instead'' or ``save
the object to disk''. One of the authors of this paper manually writes cells
that check if a result file exists, and perform the calculation only if it's not
there.

%% Rick on auditing (possible solution): ``you didn't see your mistakes, they were
%% in there but you never really saw them because a mistake never led to an answer
%% at the end'' ``the real result of this analysis was these three plots, so go
%% back and figure out everything that I did that was involved in creating those
%% three plots. So that I could start from the raw data and create those three
%% plots.''

@@

Taylor, who writes packages and example notebooks for them, seems to be
haunted the most by dead notebooks.  At first, he had to learn how to write safe
notebooks: ``It would be a file inside a directory that only I had [...] access
to and so someone else couldn't run it, [...] a temporary file on something that
was read-only to everyone else. But I've gotten careful about that.''

But even correcting notebooks can come too late: ``I'll go back and change it and update
that page, but then what happens is people have in the meantime forked it. [...]
Usually what it'll be is that I'll have to change a package as well, so
their old fork stops working, and then they complain.''

Taylor tried to keep his notebook tree tidy, but this didn't help, because
people either keep old links in their email, or forked a notebook which is now
obsolete. By design, notebooks that are deleted in the user interface are not
purged from the repository. ``Now I'm really gun-shy about sharing notebooks
because it's like, do I want to support this forever?''

Taylor and Horace work together on a notebook:
\begin{quote}
There's no way for both of us
to have ownership of a notebook, so the only way is to fork it back and forth,
and so we have dozens of old copies. [We end up] deleting all the old ones, but
people still have links to them, because they don't actually disappear. [...]
The UI makes it easy to go back to the old stale versions, because you can say
`forked from' `forked from' `forked from', but there's actually no way to go the
way that we want people to go, which is up the tree.
\end{quote}
